---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme_figures/README-"
)
```

# multilayer.ergm: Exponential Random Graph Models for Multilayer Networks

## Updates
Version 0.2.2 has been released. This version improves performance and updates dependencies with `statnet` packages.

Version 0.2.1 has been released. This version adds the ability to specify `attr` for the `duplexdyad` multilayer term.

Version 0.2.0 has been released. This version updates the interface with the `statnet` packages.

Version 0.1.6 has been released. This version ensures reproduction of the results in the package's companion paper published in _Political Science Research and Method_.

## About 

This package extends the exponential random graph model (ERGM) to multilayer networks. For more information on the application of ERGMs to multilayer networks, see [Chen, 2021, "Statistical Inference for Multilayer Networks in Political Science." _Political Science Research and Methods_ 9(2)](https://doi.org/10.1017/psrm.2019.49).

## Installation

To install the package from [GitHub](https://github.com/tedhchen/multilayer.ergm), you will need the `devtools` package. If you are on a Windows machine, you will also need the standalone [Rtools](https://cran.r-project.org/bin/windows/Rtools/). If you do not already have the `network`, `ergm`, and `statnet.common` packages, installing `multilayer.ergm` will install them as well. 

Install the package by entering the following into your R console. The argument ``build_opts`` can be removed entirely if you do not require the package tutorial. 
```{r, eval=F}
devtools::install_github("tedhchen/multilayer.ergm", build_opts = c("--no-resave-data", "--no-manual"))
```

Once installed, the package tutorial can be accessed by entering ```vignette('policy_multiplex')``` into your R console. The tutorial can also be accessed at [my website](https://tedhchen.com/pages/policy_multiplex.html).



